Generalized Tractability

### 2023-04-08 GeneralizedTractability_230310.tex
Updates from Overleaf

\newcommand{\fred}[1]{\begingroup\color{blue}Fred: #1\endgroup}
\newcommand{\peter}[1]{\begingroup\color{purple}#1\endgroup}
\newcommand{\kachi}[1]{\begingroup\color{ForestGreen}#1\endgroup}

@@

Note that the tractability function is bounded below away from zero even as $\varepsilon \to \infty$, so that the information based complexity is not expected to vanish under arbitrarily loose error requirements.  For this definition of tractability to make sense we assume that
\begin{subequations} \label{eq:Tconditions}
\begin{equation}
	T \text{ is non-decreasing in all variables}
\end{equation}
which implies that the problem does not become easier by decreasing the tolerance, $\varepsilon$, or increasing the dimension, $d$. Furthermore, increasing $\vp$ does not decrease $T(\cdot, \cdot, \vp)$.  Since there are an infinite number of positive eigenvalues, it also make sense to assume that
\begin{equation}
	\lim_{\varepsilon \to 0} T(\varepsilon^{-1},d,\vp) = \infty \qquad \forall d \in \naturals, \ \vp \in [0,\infty)^s.
\end{equation}
Since $T(\cdot,d,\vp)$ is non-decreasing, we may define the following limit:
\begin{equation}
	T(0,d,\vp):=\lim_{\varepsilon\to\infty}T(\varepsilon^{-1},d,\vp).
\end{equation}
Finally, there is a technical assumption required for our analysis.  There exists a $K_{\vp,\tau}$ depending on $\vp$ and $\tau$, but  independent of $\varepsilon$ and $d$, such that
\begin{equation} \label{eq:ptauassume}
	[T(\varepsilon^{-1},d,\vp)]^\tau \le K_{\vp,\tau} T(\varepsilon^{-1},d,\tau \vp),   \quad \forall \varepsilon \in (0,\infty), \ d \in \naturals, \ \vp\in[0,\infty)^s, \ \tau\in [1,\infty).
\end{equation}
\end{subequations}

@@

Note now that, for any $d\in\naturals$,
\begin{align*}
 i(\infty,d,\vp') & = \lfloor C_{\vp'}\, T(0,d,\vp')\rfloor +1 \le  \left[C_{\vp'} + \frac{1}{T(0,d,\vp')}   \right]\, T(0,d,\vp')\\
 & \le  \underbrace{\left[C_{\vp'} + \frac{1}{T(0,1,\vp')}   \right]}_{=:L_{\vp'}}\, T(0,d,\vp')
 \le \lceil L_{\vp'} T(0,d,\vp')\rceil .
\end{align*}
For this choice of $ L_{\vp'}$ we have
\[
\sup_{d\in\naturals}}\,\, \sum_{i=\lceil L(\vp') T(0,d,\vp')\rceil}^\infty \frac{1}{T(\lambda_{i,d}^{-1},d, \vp')}
\le
\sup_{d\in\naturals} \, \, \sum_{i=i (\infty,d,\vp')+1}^\infty \frac{1}{T (\lambda_{i,d}^{-1},d,\vp')} <
\infty.
\]
This yields \eqref{eq:tractiff4} with $\vp$ replaced by $\vp'$, so we see the necessity of $\eqref{eq:tractiff4}$.

@@

\fred{We need to check the proof for
\begin{itemize}
\item correctness
\item exponents
\item whether the additional assumptions are too restrictive and can be generalized
\item whether we need $\tau$ in the assumption
\item Do we need non-decreasing or strictly increasing?
\end{itemize}}
\begin{definition}[Weak Tractability]
If $n(\varepsilon,d)$ denotes the minimal number of function values needed to approximate all functions. We say that the problem is \emph{weakly tractable} if
    \[\lim_{\varepsilon^{-1}+d\rightarrow \infty} \frac{\ln n(\varepsilon,d)}{\varepsilon^{-1}+d} = 0\]
\end{definition}
\begin{definition}[Quasi-polynomial Tractability]
    The problem is \emph{quasi-polynomially tractable}, if there exist two constants $C, t> 0$ such that
    \[
    n(\varepsilon,d) \leq C\exp\{t(1+\ln(1/\varepsilon))(1+\ln(d))\}
    \] for all $\varepsilon \in (0,1)$ and $d\in\mathbb{N}$.
\end{definition}

@@

We show a necessary and sufficient condition for $T$-$(s,t)$-weak tractability, working analogously to \cite{WW17}.
\begin{theorem}\label{thm:weak}
For $c>0$, define
\[
\sigma (c,d,s,\vp):= \sum_{i=1}^\infty \exp \left(-c\, T(\lambda_{i,d}^{-s},1,\vp)\right).
\]
Then the problem is $T$-$(s,t)$-weakly tractable if and only if
\[
\mu (c,s,t,\vp)=\sup_{d\in\naturals} \sigma (c,d,s,\vp) \exp\left(-c\, T(0,d^t,\vp)\right)<\infty
\]
holds for all $c>0$.
\end{theorem}
\begin{proof}
 We first show the necessity of the condition in the theorem.   Indeed, suppose that the problem is  $T$-$(s,t)$-weakly tractable. Then, for any $c>0$ there exists a positive integer $V=V(c)$ such that  
 \[
 n(\varepsilon,d)\le 
 \left\lfloor \exp\left(c\,
 \left(T(\varepsilon^{-s},1,\vp) + T(0,d^t,\vp)\right)
 \right) \right\rfloor
 \]
 whenever $\varepsilon^{-1} + d \ge V$. For $d\in\naturals$ 
 choose $\varepsilon>0$ such that $\varepsilon^{-1}\ge \max\{1,V-d\}$. Since the eigenvalues $\lambda{i,d}$ are non-increasing in $i$, we know that
 \[
 \lambda_{\lfloor \exp(c\,
 (T(\varepsilon^{-s},1,\vp) + T(0,d^t,\vp))) \rfloor+1}^2 \le \varepsilon^2
 \]
 for any $\varepsilon$ satisfying $\varepsilon^{-1}\ge \max\{1,V-d\}$.
Let 
\[
n=n(\varepsilon):= \lfloor \exp(c\,
 (T(\varepsilon^{-s},1,\vp) + T(0,d^t,\vp))) \rfloor+1
\]
and
\[
n^*:= \lfloor \exp(c\,
 (T(\max\{1,(V-d)^s\},1,\vp) + T(0,d^t,\vp))) \rfloor+1.
\]
By varying $\varepsilon^{-1}$ ini the interval 
$[\max\{1,V-d\},\infty)$ we can make $n(\varepsilon)$ attain 
any integer value greater than or equal to $n^*$.

@@

\end{proof}
}

>>>>

\newcommand{\fred}[1]{\begingroup\color{blue}#1\endgroup}
\newcommand{\peter}[1]{\begingroup\color{purple}#1\endgroup}
\newcommand{\kachi}[1]{\begingroup\color{ForestGreen}#1\endgroup}

@@

Note that the tractability function is bounded below away from zero even as $\varepsilon \to \infty$, so that the information-based complexity is not expected to vanish under arbitrarily loose error requirements.  For this definition of tractability to make sense we assume that
\begin{subequations} \label{eq:Tconditions}
\begin{equation}
	T \text{ is non-decreasing in all variables}
\end{equation}
which implies that the problem does not become easier by decreasing the tolerance, $\varepsilon$, or increasing the dimension, $d$. Furthermore, increasing $\vp$ does not decrease $T(\cdot, \cdot, \vp)$.  Since there are an infinite number of positive eigenvalues, it also make sense to assume that
\begin{equation}
	\lim_{\varepsilon \to 0} T(\varepsilon^{-1},d,\vp) = \infty \qquad \forall d \in \naturals, \ \vp \in [0,\infty)^s.
\end{equation}
Since $T(\cdot,d,\vp)$ is non-decreasing, we may define the following limit:
\begin{equation}
	T(0,d,\vp):=\lim_{\varepsilon\to\infty}T(\varepsilon^{-1},d,\vp).
\end{equation}
Finally, there is a technical assumption required for our analysis.  There exists a $K_{\vp,\tau}$ depending on $\vp$ and $\tau$, but  independent of $\varepsilon$ and $d$, such that
\begin{equation} \label{eq:ptauassume}
	[T(\varepsilon^{-1},d,\vp)]^\tau \le K_{\vp,\tau} T(\varepsilon^{-1},d,\tau \vp),   \quad \forall \varepsilon \in (0,\infty), \ d \in \naturals, \ \vp\in[0,\infty)^s, \ \tau\in [1,\infty).
\end{equation}
\end{subequations}

@@

Note now that, for any $d\in\naturals$,
\begin{align*}
 i(\infty,d,\vp') +1  & = \lfloor C_{\vp'}\, T(0,d,\vp')\rfloor +2 \le  \left[C_{\vp'} + \frac{1}{T(0,d,\vp')}   \right]\, T(0,d,\vp')\\
 & \le  \underbrace{\left[C_{\vp'} + \frac{2}{T(0,1,\vp')}   \right]}_{=:L_{\vp'}}\, T(0,d,\vp')
 \le \lceil L_{\vp'} T(0,d,\vp')\rceil .
\end{align*}
For this choice of $L_{\vp'}$ we have
\[
\sup_{d\in\naturals}\,\, \sum_{i=\lceil L(\vp') T(0,d,\vp')\rceil}^\infty \frac{1}{T(\lambda_{i,d}^{-1},d, \vp')}
\le
\sup_{d\in\naturals} \, \, \sum_{i=i (\infty,d,\vp')+1}^\infty \frac{1}{T (\lambda_{i,d}^{-1},d,\vp')} <
\infty.
\]
This yields \eqref{eq:tractiff4} with $\vp$ replaced by $\vp'$, so we see the necessity of $\eqref{eq:tractiff4}$.

@@

\fred{We need to check the proof for
\begin{itemize}
\item correctness
\item exponents
\item whether the additional assumptions are too restrictive and can be generalized
\item whether we need $\tau$ in the assumption
\item Do we need non-decreasing or strictly increasing?
\end{itemize}}
\begin{definition}[Weak Tractability]
If $n(\varepsilon,d)$ denotes the minimal number of function values needed to approximate all functions. We say that the problem is \emph{weakly tractable} if
    \[\lim_{\varepsilon^{-1}+d\rightarrow \infty} \frac{\ln n(\varepsilon,d)}{\varepsilon^{-1}+d} = 0\]
\end{definition}
\begin{definition}[Quasi-polynomial Tractability]
    The problem is \emph{quasi-polynomially tractable}, if there exist two constants $C, t> 0$ such that
    \[
    n(\varepsilon,d) \leq C\exp\{t(1+\ln(\varepsilon^{-1}))(1+\ln(d))\}
    \] for all $\varepsilon \in (0,1)$ and $d\in\mathbb{N}$.
\end{definition}

@@

We show a necessary and sufficient condition for $T$-$(s,t)$-weak tractability, working analogously to \cite{WW17}.
\begin{theorem}\label{thm:weak}
For $c>0$, define
\[
\sigma (c,d,s,\vp):= \sum_{i=1}^\infty \exp \left(-c\, T(\lambda_{i,d}^{-s},1,\vp)\right).
\]
Then the problem is $T$-$(s,t)$-weakly tractable if and only if
\[
\mu (c,s,t,\vp)=\sup_{d\in\naturals} \sigma (c,d,s,\vp) \exp\left(-c\, T(0,d^t,\vp)\right)<\infty
\]
holds for all $c>0$.
\end{theorem}
\begin{proof}
 We first show the necessity of the condition in the theorem.   Indeed, suppose that the problem is  $T$-$(s,t)$-weakly tractable. Then, for any $c>0$ there exists a positive integer $V=V(c)$ such that  
 \[
 n(\varepsilon,d)\le 
 \left\lfloor \exp\left(c\,
 \left(T(\varepsilon^{-s},1,\vp) + T(0,d^t,\vp)\right)
 \right) \right\rfloor
 \]
 whenever $\varepsilon^{-1} + d \ge V$. For $d\in\naturals$ 
 choose $\varepsilon>0$ such that $\varepsilon^{-1}\ge \max\{1,V-d\}$. Since the eigenvalues $\lambda{i,d}$ are non-increasing in $i$, we know that
 \[
 \lambda_{\lfloor \exp(c\,
 (T(\varepsilon^{-s},1,\vp) + T(0,d^t,\vp))) \rfloor+1}^2 \le \varepsilon^2
 \]
 for any $\varepsilon$ satisfying $\varepsilon^{-1}\ge \max\{1,V-d\}$.
Let 
\[
n=n(\varepsilon):= \lfloor \exp(c\,
 (T(\varepsilon^{-s},1,\vp) + T(0,d^t,\vp))) \rfloor+1
\]
and
\[
n^*:= \lfloor \exp(c\,
 (T(\max\{1,(V-d)^s\},1,\vp) + T(0,d^t,\vp))) \rfloor+1.
\]
By varying $\varepsilon^{-1}$ in the interval 
$[\max\{1,V-d\},\infty)$ we can make $n(\varepsilon)$ attain 
any integer value greater than or equal to $n^*$.

@@

\end{proof}
}
\paragraph{Example 3: EXP-strong polynomial tractability}

Let $T: (0,\infty) \times \naturals \times [0,\infty) \times [0,\infty)\Rightarrow [1,\infty)$
be defined by
\[
 T(\varepsilon^{-1},d,\vp)= (1+\log\max\{\varepsilon^{-1},1\})^p.
\]
Then Theorem \ref{thm_main_tract2} yields the iff-condition
\[
 \sup_{d\in\naturals} \sum_{i=L_{\vp} d^{q_1}}^\infty (\min\{\lambda_{i,d},1\})^p  < \infty.
\]
\paragraph{Example 4: EXP-polynomial tractability}

Let $T: (0,\infty) \times \naturals \times [0,\infty) \times [0,\infty)\Rightarrow [1,\infty)$
be defined by
\[
 T(\varepsilon^{-1},d,\vp)= (1+\log \max\{\varepsilon^{-1},1\})^p\, d^q.
\]
Then Theorem \ref{thm_main_tract2} yields the iff-condition
\[
 \sup_{d\in\naturals} \sum_{i=L_{\vp} d^{q_1}}^\infty (\min\{\lambda_{i,d},1\})^p d^{-q} < \infty.
\]
\paragraph{Example 5: $(t,s)$-Weak tractability}

Let $T: (0,\infty) \times \naturals \times [0,\infty) \times [0,\infty)\Rightarrow [1,\infty)$. Suppose we have $T$-$(s,t)$- weak tractability, such that
\[
 \lim_{d+\varepsilon^{-1}\rightarrow\infty} \frac{\log \max(1,n(\varepsilon,d))}{d^t +(\max(1,\varepsilon^{-1}))^s} =0.
\]
Then Theorem \ref{thm:weak} yields the iff-condition
\[
\sup_{d\in\naturals} \sigma (c,d,s,\vp) \exp\left(-c\, T(0,d^t,\vp)\right)<\infty
\]
where $\sigma (c,d,s,\vp) = $

\section{$T$-Quasi-polynomial tractability} \label{sec:qpt}

\begin{definition}[$T$-Quasi-polynomial tractability]
We say that the problem is \emph{$T$- quasi-polynomial tractable} if there exists positive $C_{\vp}$ such that
   %\[ n(\varepsilon,d)\leq C_{\vp}\exp\{(T(0,d,\vp))(T(\varepsilon^{-1},1,\vp)\}\]
   \[ n(\varepsilon,d)\leq C_{\vp}\exp\{(T(\varepsilon^{-1},d,\vp)\}\]
\end{definition}

\begin{theorem}\label{thm:qpt}
A problem is $T$-quasi-polynomial tractable iff there exists $\vp \in [\vzero, \boldsymbol{\infty})$ such that
\begin{equation}\label{thmqpt}
M_{\vp} := \sup_{d\in\mathbb{N}}\sum_{i=1}^\infty \frac{1}{T(\lambda_{i,d}^{-1},d,\vp)}
<\infty
\end{equation}
%where $\Bar{p} = -p(1+\log(T(0,d,\vp))$.
If it holds then 
\[n(\varepsilon,d)\leq 1+T(0,d^\tau,\vp)+ T(0,d^\tau,\vp)T(\varepsilon^{-1},1,\vp)\]
\end{theorem}
\begin{proof}
\textbf{Necessity Condition:}\\
Suppose that we have $T$-quasi-polynomial tractability, i.e., for some $\vp\geq 0$, there exists positive $C_{\vp}$ such that
%\[\comp(\varepsilon,d)\le C_{\vp}\exp\{(T(0,d,\vp)(T(\varepsilon^{-1},1,\vp))\}\quad\quad \forall\varepsilon >0, d\in \mathbb{N}\] 
\[
\comp(\varepsilon,d)\le C_{\vp}\exp\{T(\varepsilon^{-1},d,\vp)\}
\quad\quad \forall\varepsilon >0, d\in \mathbb{N}
\] 
Since $\lambda_{i,d}$ are non-increasing, we have
\[
\lambda_{\left\lfloor C_{\vp}\exp\{T(\varepsilon^{-1},d,\vp)\}\right\rfloor+1,d}\leq \varepsilon^2
\]
Given that the estimate holds for all $\varepsilon> 0$, we assume that $\varepsilon \in (0, 1]$.\\

If we vary $\varepsilon\in (0,1]$, we have that
\[
j(\varepsilon,d,\vp) = \left\lfloor C_{\vp}\exp\{T(\varepsilon^{-1},d,\vp)\}\right\rfloor+1
\] which attains the values $j = j_d,j_{d+1},\dots$, where
\[
j(\infty,d,\vp)=j_d = \lfloor C_{\vp}\exp\{ T(0,d,\vp)\}\rfloor + 1
\]
Furthermore, we have
\[j \leq C_{\vp}\exp\{T(\varepsilon^{-1},d,\vp)\}+1 \leq C_{\vp}\exp\{T(\lambda^{-1}_{j,d},d,\vp)\}+1 \]
This implies that
\[
\left(\log\frac{j-1}{C_{\vp}}\right)^\tau \leq \left(T(\varepsilon^{-1},d,\vp)\right)^\tau \leq K_{\vp,\tau}T(\lambda_{j,d},d,\tau\vp) 
\]
we have 
%\[\frac{1}{T(\lambda_{i,d}^{-1},d,\vp)} \leq \frac{K_{\vp,\tau}}{\left(\log\frac{j-1}{C_{\vp}}\right)^\tau}\]
\end{proof}
